### 2016-09-01 doc/manuscript.Rmd
Added an onload function to load up settings/data
Mid-state of manuscript writing (at least to have a baseline in history)

@@

- Primary analysis:
    - Heatmap
    - GEE
        - Include # of both uncorrected and corrected (FDR) significant associations
        - QIC modeling
        - DAG development
        - Look over the entire model (see which variables are significant, etc)
    - Interaction (manhattan plot)
    - Interaction plots (with total? + clinical?)
        - With Ethnicity, generate contrast plots of adjusted estimates
    - PLS with ISI and HOMA-IR
        - CV for determining number of comps
        - Loadings (of TAGFA)
        - Prediction (of predicted vs real)
        - Use extracted PLS components in GEE
        - Train and test (CV) to see how it predicts in a new dataset (ie. a future visit)
- Secondary analysis:
    - PLS-DA (dysgly)
        - Train and test (CV)
        - Loadings (of TAGFA) 
        - Groupings
        - Prediction (misclassification)
    - AROC
        - dysgly with clinical TAG vs the 'best' fatty acid to compare to

@@

@@

```{r load_data, message=FALSE}
devtools::load_all('.')
set_options()
load_data()
ds <- project_data
```

@@

fig_gee <- fig('gee', 'caption here')
```

```{r set_datasets}
pls_df_homa <- prep_pls_data(ds, 'linvHOMA')
pls_df_isi <- prep_pls_data(ds, 'lISI')
```

```{r analyze_gee, cache=TRUE}
gee_results <- analyze_gee(ds)
gee_results_int <- analyze_gee(ds, intvar = 'VN')
```

```{r analyze_pls}
pls_results <- list()
pls_results$homa$train <- analyze_pls(pls_df_homa$train, 'linvHOMA')
pls_results$homa$test <- analyze_pls(pls_df_homa$test, 'linvHOMA')
pls_results$isi$train <- analyze_pls(pls_df_isi$train, 'lISI')
pls_results$isi$test <- analyze_pls(pls_df_isi$test, 'lISI')
```

```{r analyze_cor}
cor_results <- analyze_corr(ds)
```

@@

## GEE

@@

Interaction effect

## PLS

@@

## PLS-DA

Misclassification of dysgly {{using test?}}, misclass by group too ('yes' vs 'no')

## AROC

AROC of clinical TAG vs TAGFA palmitic acid etc.

@@

@@

```{r fig_heatmap}
plot_heatmap(cor_results)
```

```{r fig_dist}

```

```{r table_dist}
table_distribution()
```

@@

```{r fig_gee}
plot_gee_main(gee_results)
```

```{r table_gee_main}
table_gee_main(gee_results, '')
```

```{r fig_gee_int}
seer::view_interaction(gee_results_int, 
                       groups = 'unit~Yterms',
                       ylab = 'Triacylglycerol fatty acids') +
        graph_theme()
```

```{r fig_interacts}
# plot_interactions(ds, fa, outcome)
# plot_interactions(ds, fa, outcome)
# plot_interactions(ds, fa, outcome)
```

```{r fig_pls_cvncomp}
# Or as a table?
plot_pls_cv_ncomp(pls_results$homa$train)
plot_pls_cv_ncomp(pls_results_homa)
```

```{r fig_pls_loadings}
# As a single figure
seer::view_pls_xloadings(pls_results$homa$train) +
    graph_theme()
seer::view_pls_xloadings(pls_results$homa$test) +
    graph_theme()
seer::view_pls_xloadings(pls_results$isi$train) +
    graph_theme()
seer::view_pls_xloadings(pls_results$isi$test) +
    graph_theme()
```

```{r fig_pls_prediction}
# maybe not a figure, maybe just a correlation?
predplot(pls_results$isi$train, ncomp = 1:2)
predplot(pls_results$homa$train, ncomp = 1:2)
plot_pls_predict(pls_results$isi$train)
plot_pls_predict(pls_results_homa)
plot_pls_predict(test_set, pls_predict_set)
```

```{r fig_pls_comp_gee}
plot_gee_main(gee_results_plscomps)
```

```{r table_gee_pls}
table_gee_main(gee_results_plscomps)
```

```{r fig_plsda_cvncomps}
# can't remember if this is possible...
plot_plsda_cv_ncomp(plsda_results)
```

```{r fig_plsda_loadings}
plot_plsda_loadings(plsda_results)
```

```{r fig_plsda_groupings}
plot_plsda_groupings(plsda_results)
```

```{r fig_aroc}
plot_aroc(aroc_results) #?
```

>>>>

Note: keep the story simple and clean, without getting too dense

@@

- Primary analysis:
    - Heatmap
    - GEE
        - Include # of both uncorrected and corrected (FDR) significant associations
        - QIC modeling
        - DAG development
        - Look over the entire model (see which variables are significant, etc)
    - PLS with ISI and HOMA-IR
        - CV for determining number of comps (best is 2)
        - Loadings (of TAGFA) (tho all are the same, use HOMA/ISI. BCF have
        lower corr between measured and predicted) (Use COrr of both components
        (1 and 2))
        - Prediction (of predicted vs real) (in the pls.rmd file, not in
        manuscript (or maybe in ESM?))
        - Use extracted PLS components in GEE (?)
        - Train and test (CV) to see how it predicts in a new dataset (ie. a
        future visit) (done. In pls.rmd file. As ESM?)
- Secondary analysis:
    - PLS-DA (dysgly)
        - Train and test (CV)
        - Loadings (of TAGFA) 
        - Groupings
        - Prediction (misclassification)
    - AROC
        - dysgly with clinical TAG vs the 'best' fatty acid to compare to

@@

NEFA may be a mediating factor in association, but that adjustment will not
attenuate this assoc since TAGFA composition is also dictated by other sources
for DNL (e.g. recent meal, DNL from glucose stores, etc.).

@@

```{r load_data, message=FALSE}
devtools::load_all('.')
ds0yr <- project_data %>% 
    dplyr::filter(VN == 0)
```

@@

fig_gee <- fig('gee', 'caption here')
fig_consort <- fig('consort', 'caption here')
```

```{r analyze_gee, cache=TRUE}
gee_df <- analyze_gee()
```

```{r analyze_others}
# CV results are shown in pls.Rmd vignette
pls_homa <- analyze_pls(y = 'linvHOMA', cv = FALSE)
pls_isi <- analyze_pls(y = 'lISI', cv = FALSE)

# Correlation between predicted and actual
pls_homa_t <- analyze_pls(y = 'linvHOMA')
pls_isi_t <- analyze_pls(y = 'lISI')

cor_df <- analyze_corr()
```

Participants from London and Toronto, Canada, were recruited into the 
Prospective Metabolism and Islet cell Evaluation (PROMISE) cohort. Eligibility 
for recruitment into PROMISE required having one or more risk factors for type 2
diabetes mellitus, including obesity, hypertension, family history of diabetes, 
and/or a history of gestational diabetes or birth of a macrosomic infant. 
Participants aged 30 years and older (n=736) attended the baseline visit
between 2004-2006. Follow-up examinations in this cohort occur every three years, with three
examination visits completed to date (2004-2006, 2007-2009, and 2010-2013). 
Participants are contacted annually by telephone. The
current study sample used data on participants who did not have diabetes at baseline,
who returned for one or more of the follow-up examinations, and who had samples
available for fatty acid measurements (n=477). A diagram of the sample size at
each visit is shown in `r cite_sf('consort')`. At each examination,
participants undergo metabolic characterization, anthropometric measurements,
and questionnaires on lifestyle and sociodemographics. Research ethics approval
was obtained from Mount Sinai Hospital and the University of Western Ontario,
and all participants provided written informed consent.
Data collection methods were standardized across the 2 centres and research
nurses were centrally trained.

## Blood measure assessments

At each examination, an 8-12 hour fasting blood sample was drawn from each
participant, followed by a 75g oral glucose tolerance test (OGTT) with a 30
minute and 2 hour blood draw. All blood samples were processed and frozen at
-70°C. Alanine aminotransferase (ALT) was measured using standard laboratory
procedures. Cholesterol, HDL, and clinically-measured triacylglycerides (TAG)
were measured using Roche Modular's enzymatic colorimetric tests (Mississauga,
ON).  Both insulin and glucose were derived from the OGTT at fasting, 30 minute,
and 2 hour time points. Specific insulin was measured using the Elecsys 1010
(Roche Diagnostics, Basel, Switzerland) immunoassay analyzer and
electrochemiluminescence immunoassay. This assay shows 0.05% cross-reactivity to
intact human pro-insulin and the Des 31,32 circulating split form (Linco Res.
Inc), and has a coefficient of variation (CV) of 9.3%. Glucose was determined
using an enzymatic hexokinase (Roche Modular, Roche Diagnostics) with a
detection range of 0.11 (2 mg/dL) to 41.6 mmol/L.  The inter-assay %CV is <1.1%
and intra-assay %CV is < 1.9%.  All assays were performed at the Banting and
Best Diabetes Centre Core Lab at Mt Sinai Hospital. Impaired fasting glucose
(IFG), impaired glucose tolerance (IGT), and diabetes were categorized using the
2006 WHO criteria [@WHO2006].

TAGFA composition was quantified using stored fasting serum samples from the 
baseline visit, which had been frozen at -70°C for 4-6 years and had not been 
exposed to any freeze-thaw cycles. Serum fatty acids have been documented to be 
stable at these temperatures for up to 10 years [@Matthan2010a]. A known amount 
of heptadecanoic acid (17:0) was added to the serum as an internal standard
prior to extracting total lipids according to the method of Folch [@Folch1957a].
Each serum lipid fraction---non-esterified fatty acids (NEFA), cholesteryl
ester, phospholipid, and TAGFA---was isolated using thin layer chromatography;
each fraction was visualized under UV light after lightly spraying with
8-anilino-1-naphthalene sulfonic acid (0.1% wt/vol) and then converted to fatty
acid methyl esters with 14% boron trifluoride in methanol at 100°C for 1 h.
Fatty acid methyl esters were separated and quantified using a Varian-430 gas
chromatograph (Varian, Lake Forest, CA, USA) equipped with a Varian Factor Four
capillary column and a flame ionization detector, which were injected in
splitless mode. Fatty acid concentrations (nmol/ml) were calculated by
proportional comparison of gas chromatography peak areas to that of the internal
standards [@Nishi2014a]. There were 22 fatty acids measured in the TAGFA
fraction. Given their diverse biology, as well as the complexity of the
analyses, findings for other lipid fractions in this cohort are reported
separately (see ref @Johnston2016a and for the analysis of the phospholipid and
cholesteryl ester fractions and {{next citation}} for the analysis of the NEFA
fraction).

## Outcome variables

Insulin sensitivity and beta-cell function indices were computed using the OGTT 
glucose and insulin data. Insulin sensitivity was assessed using 1/HOMA-IR (1
divided by HOMA-IR) [@Matthews1985] and the Insulin Sensitivity Index (ISI)
[@Matsuda1999]. HOMA-IR largely reflects hepatic insulin resistance, while ISI
reflects whole-body insulin sensitivity [@AbdulGhani2007].  Beta-cell function
was assessed using the Insulinogenic Index [@Wareham1995] over HOMA-IR (IGI/IR)
and the Insulin Secretion-Sensitivity Index-2 (ISSI-2) [@Retnakaran2009]. IGI/IR
is a measure of the first phase of insulin secretion while ISSI-2 is analogous to
the disposition index (but using OGTT values). Each index has been validated
against gold standard measures [@Matthews1985; @Matsuda1999; @Retnakaran2009].

## Anthropometrics and sociodemographics

Height, weight, and waist circumference (WC) were measured at all clinic 
examinations using standard procedures.  WC was measured at the natural waist, 
defined as the narrowest part of the torso between the umbilicus and the xiphoid
process. BMI was calculated by dividing weight (kg) by height (m) squared.
Sociodemographic information, including age, sex, and ethnicity, were determined
using questionnaires administered at each examination. In the lifestyle
questionnaire, physical activity was determined using a version of the
Modifiable Activity Questionnaire (MAQ) [@Kriska1990].  The MAQ collects 
information on leisure and occupational activity, including intensity, 
frequency, and duration, over the past year.  Each reported activity from the 
MAQ was weighted by its metabolic intensity allowing for the estimation of 
MET-hours per week.

## Statistical analysis

The primary outcome variables for this analysis were 
1/HOMA-IR, ISI, IGI/IR, and ISSI-2, which were log-transformed for
the statistical modeling. The primary predictor variables for this analysis were
22 individual TAGFA as mole percent (mol%) of the total fraction and as
concentration (nmol/mL). Pearson correlation coefficients were computed to
assess the relationships of individual TAGFA with other continuous variables.

Of note is clinically measured TAG was only quantified on a portion of the
6-year visit (n = ~120 {{ find true number }}).

For the primary analysis, generalized estimating equation (GEE) models 
[@Zeger1986a] were used to determine the longitudinal associations between the 
outcome variables and the predictor variables. Given the longitudinal design 
an auto-regressive of order 1 (AR1) correlation matrix was chosen for the GEE models, 
though other matrices (eg. exchangeable) had similar fit (data not shown). GEE
is well suited to longitudinal cohort studies given it's capacity to handle
missed visits. The predictor variables and continuous covariates were scaled
(mean centered and standardized). The TAGFA variables were classified as
*time-independent* (held constant) as they were measured only at the baseline
visit, while the outcome variables and covariates were set as *time-dependent*.
No imputation was conducted on missing values.

Covariate selection was based on previous literature, directed acyclic graph 
[@Greenland1999a] recommendations, and quasi-likelihood information criterion 
(QIC) {{citation?}}. `r cite_st('qic')` shows the covariates compared using QIC. While the
final GEE model selected as best fitting differed between insulin sensitivity
and beta-cell function measures, the differences in QIC values were less than
10 between the models, suggesting similar fit. As such, we selected the model
that had the fewest covariates and that had similar fit between the outcome
measures ({{##}}). The final GEE model was adjusted for {{covariates}}. After scaling,
log-transforming, and exponentiating, the GEE estimates are
interpreted as an expected percent difference in the outcome variable for every 
standard deviation (SD) increase in the predictor variable given the covariates 
are held constant. Lastly for the GEE models, we tested for an interaction with
sex, ethnicity, or time by the predictor term for each outcome variable.

While GEE accounts for the longitudinal design of the data, this approach is 
limited in that it cannot analyze the inherent multivariate nature of the 
composition of the TAGFA fraction. To confirm the GEE results in a multivariate 
environment (all TAGFA), partial least squares regression (PLS) {{citation}} was used to
identify the patterns of TAGFA composition against insulin sensitivity and
beta-cell function. For more detailed explanation of GEE and PLS, please see the
{{supplemental}} Methods.

All analyses were performed using R `r getRversion()` [@Rbase], along with the R
packages geepack `r packageVersion('geepack')` for GEE [@Hoejsgaard2006a] and pls
`r packageVersion('pls')` for PLS {{cite?}}. The R code and extra analyses for
this manuscript is available at {{code doi}}. Results were considered
statistically significance at p<0.05, after adjusting for multiple testing using
the Benjamini-Hochberg False Discovery Rate [@Benjamini1995a]. STROBE was used
as a guideline for reporting [@Vandenbroucke2007b].

The number of components to extract from PLS was determined by 10-fold CV.

Dataset was split into a training set and a testing set to determine the
predictability of the given PLS model. Final results presented use the full dataset.a

@@

```{r inline}
dys <- project_data %>% 
    prep_dys_data() %>% 
    calc_convert_dys()

tagfa <- calc_tagfa_percent()
outcome_chg <- calc_outcome_changes()
calc_discr_npct(ds0yr$Sex, 'Female')
calc_discr_npct(ds0yr$Ethnicity, 'European')
estci <- calc_gee_estci(gee_df)
fa_cor <- calc_cor()
cor_pls_homa <- calc_pred_corr(pls_homa_t, pls_homa_t$test_data)
cor_pls_isi <- calc_pred_corr(pls_isi_t, pls_isi_t$test_data)
```

## Basic characteristics of the PROMISE cohort

Basic characteristics/demographics

The outcome variables significantly decreased over the 6 years by a median 
`r outcome_chg$chg` (p`r outcome_chg$p`; n=`r outcome_chg$n`).

`r fig_tagfa` shows the distribution of each TAGFA (see `r stab_tagfa` for mean
and SD {{ confirm }}). Three TAGFA contributed 
`r format_rounding(sum(tagfa$pct))`% to the total TAG concentration:
`r tagfa$c[1]`; `r tagfa$c[2]`; and, `r tagfa$c[3]`.

`r fig_corr` shows a heatmap of the correlation of individual TAGFA with the
outcome variables and several basic characteristics.
As expected, nearly all TAGFA had very strong positive correlations (r=
`r fa_cor$rng$TAG`) with clinically-measured TAG. There was also moderate
negative correlations with HDL (r=`r fa_cor$rng$HDL`). For the outcome variables,
the correlations for the insulin sensitivity measures were generally higher 
(1/HOMA-IR: r=`r fa_cor$rng['1/HOMA-IR']`, ISI: r=`r fa_cor$rng$ISI`) than 
for the beta-cell function measures (all r<0.30).

## Generalized estimating equation models

@@

There were {{ ## }} significant associations in the unadjusted GEE model shown
in {{ `r sfig_gee` }}. Only a few associations were attenuated after full model
adjustment primarily with the beta-cell function outcomes, as shown in 
`r fig_gee`. As a concentration, nearly all TAGFA had a strong negative
association on 1/HOMA-IR and ISI, and a few had strong negative associations
with IGI/IR and ISSI-2. As a mol%, four TAGFA (14:0, 16:0, 14:1n-7, and 16:1n-7)
had negative associations with 1/HOMA-IR and ISI (between {{ ## }}% lower
insulin sensitivity for every SD increase in the TAGFA), while several more TAGFA had
positive associations with 1/HOMA-IR and ISI (20:0, 18:1n-9, 20:1n-9, 22:1n-9,
18:2n-6, 20:2n-6, 20:4n-6, and 22:5n-3) predicting a {{ ## }}% higher insulin
sensitivity for every SD increase in the TAGFA. 
One TAGFA, 20:2n-6, had a very strong positive association with the
insulin sensitivity measures, with a {{ ## }}% higher insulin sensitivity for
every SD increase in 20:2n-6.

Both clinically-measured TAG and total TAGFA concentration had very strong
negative associations with all outcome variables.

There were no significant interactions by time, sex, or ethnicity for any of the
TAGFA (data not shown).

## Clustering of TAGFA by metabolic measures

@@

The predictive ability of the PLS analysis using the TAGFA composition

`r fig_pls`

The correlation between the predicted values of the outcome variables from the
PLS analysis compared to the actual value was `r cor_pls_homa$r` 
(p`r cor_pls_homa$p`) for 1/HOMA-IR and `r cor_pls_isi$r` (p`r cor_pls_isi$p`) 
for ISI, suggested TAGFA composition has moderate predictive ability for 
insulin sensitivity measures.

`r calc_pred_corr()`

@@

[@Rhee2011a] fasting TAG status better as discriminating between DM case vs
control than 2hr-OGTT TAG

[@Kotronen2009a] (this wasn't studied, but discussed) higher hepatic DNL may
increase amount of 'shorter' chain FA (16:0, 14:0, etc.) into VLDL and hence
into circulation. So higher of these FA may indicate higher DNL -> problems with
liver function -> maybe hepatic IR -> cyclical?

Discuss my PL and CE paper as well. Maybe also the NEFA one if it is published
by then...

The four FA that cluster together on the PLS also appear to be the same FA that
make up most of DNL, especially in context of higher carb/refined carb in diet
[@Wilke2009a].

@@

```{r fig_dist}
plot_tagfa()
```

```{r table_dist}
table_tagfa()
```

```{r fig_heatmap}
plot_heatmap(cor_df)
```

@@

```{r fig_gee}
plot_gee_main(gee_df)
```

```{r table_gee_main}
table_gee_main(gee_df)
```

```{r fig_pls}
gridExtra::grid.arrange(
    plot_pls(pls_homa) + ggplot2::ggtitle('A: 1/HOMA-IR'), 
    plot_pls(pls_isi) + ggplot2::ggtitle('B: ISI'), 
    ncol = 2)
```
